---
title: "Two Way Anova"
author: "Divit Vasu"
date: "September 11, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Two Way Anova

## Definition:

The data shows the Production of Wheat per acre for varieties of seeds and fertilizers. We set up an analysis of variance table for the following two-way results:

```{r, echo=TRUE}
df <- read.csv(file='Sample.csv')
head(df)
```

##Shapiro Test

Performing a Shapiro test on the data yields

```{r, echo=TRUE}
sh <- shapiro.test(df$Production)
sh
```

As the p-value(`r sh$p.value`) is greater than 0.05 we accept the **NULL Hypothesis** and we can say that the data for Production of wheat is normalized.

##Bartlett Test

Performing a Bartlett test on the data yields

```{r, echo=TRUE}
bartlett.test(Production~Seed, data=df)
```

Results shows that p-value is greater than 0.05 so we fail to reject the **null hypothesis** and conclude that the variances are equal across these samples.


```{r, echo=TRUE}
bartlett.test(Production~Fertilizers, data=df)
```

Results of bartlett test shows p-value greater than 0.05 for production vs. seed but for production vs. fertilizers, p-value is less than 0.05.Then also,we can perform the two way Anova Test.

##2-Way Annova on the samples
```{r, echo=TRUE}

Res.anova <- aov(Production~Seed+Fertilizers , data = df )
Res.anova
```


##Pair-wise Comparison

```{r, echo=TRUE}

TK<- TukeyHSD(Res.anova, "Seed")
TK
```

##Plot
```{r, echo=TRUE}
plot(TK)
```

##Conclusion
From the above results, we find that there is no significant difference in Production of wheat in accordance with different seeds and different fertilizers.
