---
title: "Hypothesis Testing in R"
author: "Divit Vasu"
date: "7 August 2018"
output:
  html_document:
    css: style.css
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Hypothesis Testing
We are going to perform following hypothesis tests in this practical:
* One Sample t-test
* Two Sample t-test
* Paired t-test

#One sample t-test
One sample t-test is used to test whether the population mean is equal to the specified value or not.

##Assumptions:
* The population from which, the sample drawn is assumed as Normal distribution.
* The population variance \(\sigma^2\) is unknown.

##Problem - One sample t-test
The specimen of copper wires drawn form a large lot have the following breaking strength (in kg. weight):

```{r, echo=TRUE}
strength <- c(578, 572, 570, 568, 572, 578, 570, 572, 596, 544)
wire_data <- data.frame(strength)
summary(wire_data$strength)
```

Test (using Student's t-statistic) whether the mean breaking strength of the lot may be taken to be 578 kg. weight (Test at 5 per cent level of significance).

Let \(\mu\) be the mean breaking strength of copper wires.
The hypothesis testing problem is:

\(H_0:\mu=578\) against \(H_1:\mu\ne578\)

##Check the normality
Shapiro-Wilks test is used to check the normality of the data.
```{r,echo=TRUE}
shapiro.test(wire_data$strength)
```

The p-value of the Shapiro-Wilk Normality test is 0.05113 which is not less than 0.05, we fail to reject the null hypothesis about the nomality. That is the strength of wires are normally distributed.

```{r, echo=TRUE}
qqnorm(wire_data$strength) # draw qq plot
qqline(wire_data$strength) # add reference line
```

```{r, echo=TRUE}
Result1<-t.test(wire_data$strength,mu=578,alternative="two.sided")
Result1 # display the result of t-test
```

The test statistic is -1.4917 and the p-value is 0.17. As the p-value (0.17) is greater than 0.05 (level of significance) we accept the null hypothesis at 0.05 level of significance.
There is enough evidence to support the null hypothesis that the mean strength of wires may be taken as 578 kg. weight.

#Independent Sample t-test
Independent sample t-test is used to check whether there is statistically significant difference between the means in two independent groups.

##Assumptions:
Assumptions for two sample t-test are as follows:

* The two samples are independently distributed
* The population from which, the two samples drawn are Normally distributed.
* The two population variances are unknown (equal or unequal).

##Problem - Two sample t-test
Blood pressure data from two different group of patient which are using old drug and new drug respectively are given below:

```{r, echo=TRUE}
drug_old <- c(90, 95, 67, 120, 89, 92, 100, 82, 79, 85)
drug_new <- c(71, 79, 69, 98, 91, 85, 89, 75, 78, 80)
summary(drug_old)
summary(drug_new)
```

Test at 5 per cent level whether the average blood pressure is the same between the drug_old and drug_new.

Let \(\mu_1\) and \(\mu_2\) be the mean weight of two groups.
The hypothesis testing problem is:
\(H_0:\mu_1=\mu_2\) against \(H_1:\mu_1\ne\mu_2\)

##Check the normality
Shapiro-Wilks test is used to check the normality of the data.
```{r, echo=TRUE}
shapiro.test(drug_old)
shapiro.test(drug_new)
```


The p-value of the Shapiro-Wilk Normality test is 0.6729 for drug_old and 0.8612 for drug_new which are both not less than 0.05, we fail to reject the null hypothesis about the nomality. That is the weights of chickens are normally distributed.
```{r, echo=TRUE}
qqnorm(drug_old) # draw qq plot
qqline(drug_old) # add reference line
```

```{r, echo=TRUE}
qqnorm(drug_new) # draw qq plot
qqline(drug_new) # add reference line
```

```{r, echo=TRUE}
vtest <- var.test(drug_old,drug_new,alternative = "two.sided")
vtest
```

The p-value of the test is 0.2246, which is not less than 0.05, we fail to reject the null hypothesis about the equality of variances.
There is enough evidence to support the null hypothesis that the mean blood pressure of both groups of patients may be considered equal. That is new drug has no significant impact.

Apply t test now

#Paired t-test
The paired t-test is used when we have dependent samples. It is used to compare two population means in the case of dependent. Paired t-test for dependent samples is used in 'before-after' studies, or when the samples are the matched pairs.

##Assumptions:
* The two samples are dependent.
* The difference between the two samples are independently distributed.
* The difference between the two samples are normally distributed.

##Problem - Paired t-test
In a test given to two groups of students, the marks obtained were as follows:

```{r, echo=TRUE}
first_group <- c(18, 20, 36, 50, 49, 36, 34, 49, 41, 30)
second_group <- c(29, 28, 26, 35, 30, 44, 46, 25, 32, 29)
```
Examine the significance of difference between mean marks obtained by students of the above two groups. Test at five per cent level of significance.
Let \(\mu_1\) be the mean marks of first group of students and \(\mu_2\) be the mean mean marks of second group of students. The hypothesis testing problem is:
\(H_0:\mu_1-\mu_2=0\) against \(H_1:\mu_1-\mu_2\ne0\)

```{r, echo=TRUE}
summary(first_group - second_group)
sd(first_group - second_group)
```

##Check the normality
We wish to test H_0 : Distribution of the differece is normal against H_1 : Distribution of the difference is not normal.

```{r, echo=TRUE}
resks <- ks.test(first_group, second_group, "pnorm")
resks
```

The p-value of the Kolmogorov-Smirnov Test is 0.4005 which is not less than 0.05. We fail to reject the null hypothesis about the nomality.

We conclude that the distribution of the difference between the marks of students of two different groups of same class is normally distributed.

```{r, echo=TRUE}
res01 <- t.test(first_group, second_group, paired=T, alternative="two.sided")
res01
```

The p-value of the test is 0.377, which is not less than 0.05, we fail to reject the null hypothesis about the equality of means.

There is enough evidence to conclude that there is no difference in the performance of students of both groups in same exam.

#Conclusion
In this practical we learnt how to perform hypothesis testing in R. Specifically, we performed One Sample t-test, Two Samplet-test and paired t-test on various real life problems.
